Practical Hilbert space approximate Bayesian Gaussian processes for probabilistic programming

### 2020-09-07 Paper/manuscript_v2/manuscript_07_02.tex
add diagnosis example for case study I

\usepackage{natbib}
\bibliographystyle{plainnat}
\bibpunct{(}{)}{;}{a}{,}{,}
\usepackage{subfigure}
\usepackage{multirow}
\usepackage{float}
\usepackage{soul}
\usepackage{xcolor}
\graphicspath{{./images/}}

\usepackage{amssymb}
\usepackage{amsmath}	
\usepackage{mathtools}
\usepackage{amsthm}
\usepackage{bm}
\usepackage{url}
\usepackage{parskip}

@@

Figure~\ref{ch5_fig10_Posteriors_exI} shows the posteriors predictive distributions of the three models, the exact GP, the HSGP with $m=80$ basis functions and boundary factor $c=1.2$ ($L=c\cdot 1= 1.2$; see equation (\ref{eq_boundary})), and the spline model with 80 knots. The true data-generating function and the noisy observations are also plotted. The sample observations are plotted as circles and the out-of-sample or test data are plotted as crosses. The test data located at the extremes of the plot are used for assessing model extrapolation, and the test data located in the middle are used for assessing model interpolation. The posteriors of the three models, exact GP, HSGP and spline, are similar in the interpolation regions of the input space. However, when extrapolating the spline model solution clearly differs from the exact GP and HSGP models as well as the actual observations.

@@

@@

The Stan model code for the exact GP, the approximate GP and the spline models of this case study can be found online at {\small \url{ https://github.com/gabriuma/basis_functions_approach_to_GP/tree/master/Paper/Case-study_1D-Simulated-data}}\,.
%

@@

>>>>

\usepackage{natbib}
\bibliographystyle{plainnat}
\bibpunct{(}{)}{;}{a}{,}{,}
\usepackage{subfigure}
\usepackage{multirow}
\usepackage{multicol}
\usepackage{float}
\usepackage{soul}
\usepackage{xcolor}
\graphicspath{{./images/}}

\usepackage{amssymb}
\usepackage{amsmath}	
\usepackage{mathtools}
\usepackage{amsthm}
\usepackage{bm}
\usepackage{url}
\usepackage{parskip}
\usepackage{pbox}

@@

Figure~\ref{ch5_fig10_Posteriors_exI} shows the posteriors predictive distributions of the three models, the exact GP, the HSGP with $m=60$ basis functions and boundary factor $c=1.2$ ($L=c\cdot 1= 1.2$; see equation (\ref{eq_boundary})), and the spline model with 80 knots. The true data-generating function and the noisy observations are also plotted. The sample observations are plotted as circles and the out-of-sample or test data are plotted as crosses. The test data located at the extremes of the plot are used for assessing model extrapolation, and the test data located in the middle are used for assessing model interpolation. The posteriors of the three models, exact GP, HSGP and spline, are similar in the interpolation regions of the input space. However, when extrapolating the spline model solution clearly differs from the exact GP and HSGP models as well as the actual observations.

@@

In practice, as the high computational demands in some cases of exact GPs prevents their use, to diagnose the approximation as a function of $m$, as proposed in section \ref{ch5_subsec_dependency} by using the model of relationships in Figures \ref{fig6_lscale_vs_J_vs_c} and \ref{fig6_lscale_vs_J_vs_c_Matern}, is a need. The diagnosis consists in comparing the estimated length-scale $\hat{\ell}$, for a given $m$ and $c$, with that minimum length-scale $\ell^*$ brought by the model of relationships and that can be accurately approximated for those given $m$ and $c$. Thus, if $\hat{\ell} < \ell^*$, $m$ needs to increased or $c$ decreased, and this can be iteratively until we get $\hat{\ell} \geq \ell^*$. Figure \ref{ch5_fig12_diagnosis} depicts the diagnosis for this case study. In a first iteration we choose $m=20$ and $c=1.5$, which is not a feasible combination of values as $c=1.5$ is too small for $m=20$ or $m=20$ is too small for $c=1.5$. In a second iteration we increase up to $m=30$, the HSGP estimates a length-scale as $\hat{\ell}=0.12$, and the approachable minimum length-scale that the model of relationships brings is $\ell^*=0.19$. As $\hat{\ell} < \ell^*$, the approximation should not be accurate enough (see the higher interpolation error of HSGP that GP in Figure \ref{ch5_fig11_MSE_exI_inter}-left), and $m$ needs to be increased. In a third iteration we increase up to $m=40$, the HSGP model estimates a length-scale of $\hat{\ell}=0.14$, and the model of relationships brings a approachable minimum length-scale of $\ell^*=0.15$. As $\hat{\ell}$ is still less than $\ell^*$ but both are very closed, the approximation should be almost accurate enough (see Figure \ref{ch5_fig11_MSE_exI_inter}-left where the HSGP error is approaching that of the exact GP. In a fourth iteration we increase up to $m=60$, the estimated length-scale is $\hat{\ell}=0.14$ and the minimum length-scale brought by the model of relationships is $\ell^*=0.11$, then the approximation should be accurate (The error is the same than the exact GP in Figure \ref{ch5_fig11_MSE_exI_inter}-left).

\begin{figure}
\begin{tabular}{c c c c}
{\Large 1}\, \pbox{2cm}{\footnotesize $m=20$\\ $c=1.5$}\, \pbox{1cm}{$\underset{\longrightarrow}{\text{\tiny HSGP}}$\\[-1mm]}\,  {\footnotesize$\hat{\ell}=0.09$} & {\Large 2}\, \pbox{2cm}{\footnotesize $m=30$\\ $c=1.5$}\, \pbox{1cm}{$\underset{\longrightarrow}{\text{\tiny HSGP}}$\\[-1mm]}\, {\footnotesize$\hat{\ell}=0.12$} & {\Large 3}\, \pbox{2cm}{\footnotesize $m=40$\\ $c=1.5$}\, \pbox{1cm}{$\underset{\longrightarrow}{\text{\tiny HSGP}}$\\[-1mm]}\, {\footnotesize$\hat{\ell}=0.14$} & {\Large 4}\, \pbox{2cm}{\footnotesize $m=60$\\ $c=1.5$}\, \pbox{1cm}{$\underset{\longrightarrow}{\text{\tiny HSGP}}$\\[-1mm]}\, {\footnotesize$\hat{\ell}=0.14$} \\[3mm]
\includegraphics[scale=0.21, trim = 0mm 0mm 4mm 10mm, clip]{caseI_fig6_1_1_vposter.pdf} &
\includegraphics[scale=0.21, trim = 14mm 0mm 4mm 10mm, clip]{caseI_fig6_2_1_vposter.pdf} &
\includegraphics[scale=0.21, trim = 14mm 0mm 4mm 10mm, clip]{caseI_fig6_3_1_vposter.pdf} &
\includegraphics[scale=0.21, trim = 14mm 0mm 4mm 10mm, clip]{caseI_fig6_4_1_vposter.pdf} \\[1mm]
	& $\hat{\ell} < \ell^*$ & $\hat{\ell} < \ell^*$ & $\hat{\ell} > \ell^*$
\end{tabular}
\caption{Diagnosis.}
	\label{ch5_fig12_diagnosis}
\end{figure}

@@

@@

The Stan model code for the exact GP, the approximate GP and the spline models of this case study can be found online at {\small \url{ https://github.com/gabriuma/basis_functions_approach_to_GP/tree/master/Paper/Case-study_1D-Simulated-data}}\,.
